Multiresolution Neural Field Estimation

### 2011-02-28 Multiresolution_framework/ltx/manuscript.tex
minor edits

We have a clear definition and where to stimulate and how to stimulate.

@@

Creating patient specific neural models has been approach from a number of different standpoints. \dean{need to add proper citations here.} Valdes et al. (1999) used the neural mass model of Lopes Da Silva (1976) and Zetterberg et al. (1978) to fit EEG data. This approach was extended by David and Friston (2003) to coupled neural masses (using the equations of Jansen and Ritt (1995)) through a Bayesian estimation scheme dubbed Dynamic Causal Modeling (DCM). Following this work, Galka et al. (2008), Sauer et al. (2008), Daunizeau et al. (2009), and Freestone et al (2011) developed the estimation frameworks for continuum neural field equations that explain the richer dynamics of spatiotemporal neural fields.

@@

The multiresolution approximation (MRA) of the neural IDE is be obtained by decomposing both the field, $v_t(.)$, and the connectivity kernel, $w(.)$, (assuming square-integrable functions) using translations and dilations of a scaling function $\phi(\mathbf{r})$ and a mother wavelet $\psi(\mathbf{r})$. Considering a one-dimensional field, the connectivity kernel is decomposed by,

@@

To derive the state-space model, Eqs.~\eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} are substituted into Eq. \eqref{eq:DiscreteTimeModel}, then we cross-multiply by $\boldsymbol\mu\left(\mathbf r\right)$ and integrate over the space giving
\begin{align}\label{eq:DecomposedModel2} 
	&\int_{\Omega} \boldsymbol\mu  \left(r\right)\boldsymbol\mu\left(r\right)^\top  dr\mathbf{x}_{t+1}= 
	\xi\int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu\left(r\right)^\top d\mathbf{r}\mathbf{x}_t \nonumber \\
	&+T_s \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r'\right)\mathbf{x}_t 
	\, dr'd\mathbf{r}} \nonumber \\
	+& \int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
\end{align}
The simplify the model we define the matrices
\begin{align}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) dr \\
\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r\right)\mathbf{x}_t 
	\, dr'dr}
\end{align}
and substitute them into Eq.~\eqref{eq:DecomposedModel2}. Then we cross-multiply by $\mathbf{\Lambda}_{x}^{-1}$ giving
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t \\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= (T_s\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I),
\end{align}
where $\mathbf I$ is the identity matrix. The disturbance becomes 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r},
\end{equation}
which is a vector valued, zero-mean normally distributed white noise process with covariance (see \cite{Scerri2009} for proof)
\begin{equation}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into equation \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is given by
\begin{equation}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(\mathbf{r}_i - \mathbf{r}')\boldsymbol{\mu}_j(\mathbf{r}') \, d\mathbf{r}'.
\end{equation}
This completes derivation of the state-space representation of the multiresolution approximation of the neural field model. 
% \parham{Mike you can add state-space derivation using Galerkin projection here.}

>>>>

We have a clear definition and where to stimulate and how to stimulate.\\
\cut{This color is used to indicate bits we can get rid of if we dont have enough space, use the command: \textbackslash cut$\left\lbrace  \right\rbrace$}

@@

Creating patient specific neural models has been approached from a number of different standpoints. \dean{need to add proper citations here.} Valdes et al. (1999) used the neural mass model of Lopes Da Silva (1976) and Zetterberg et al. (1978) to fit EEG data. This approach was extended by David and Friston (2003) to coupled neural masses (using the equations of Jansen and Ritt (1995)) through a Bayesian estimation scheme dubbed Dynamic Causal Modeling (DCM). Following this work, Galka et al. (2008), Sauer et al. (2008), Daunizeau et al. (2009), and Freestone et al (2011) developed the estimation frameworks for continuum neural field equations that explain the richer dynamics of spatiotemporal neural fields.

@@

The multiresolution approximation (MRA) of the neural IDE is obtained by decomposing both the field, $v_t(.)$, and the connectivity kernel, $w(.)$, (assuming square-integrable functions) using translations and dilations of a scaling function $\phi(\mathbf{r})$ and a mother wavelet $\psi(\mathbf{r})$. Considering a one-dimensional field, the connectivity kernel is decomposed by,

@@

To derive the state-space model, Eqs.~\eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} are substituted into Eq. \eqref{eq:DiscreteTimeModel}, then we cross-multiply by $\boldsymbol\mu\left(\mathbf r\right)$ and integrate over the space giving
\begin{align}\label{eq:DecomposedModel2} 
	&\int_{\Omega} \boldsymbol\mu  \left(r\right)\boldsymbol\mu\left(r\right)^\top  dr\mathbf{x}_{t+1}= 
	\xi\int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu\left(r\right)^\top d\mathbf{r}\mathbf{x}_t \nonumber \\
	&+T_s \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r'\right)\mathbf{x}_t 
	\, dr'd\mathbf{r}} \nonumber \\
	+& \int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
\end{align}
In order to simplify the model we define the matrices
\begin{align}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) dr \\
\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r\right)\mathbf{x}_t 
	\, dr'dr}
\end{align}
and substitute them into Eq.~\eqref{eq:DecomposedModel2}. Then we cross-multiply by $\mathbf{\Lambda}_{x}^{-1}$ giving
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t \\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= (T_s\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I),
\end{align}
where $\mathbf I$ is the identity matrix. The disturbance becomes 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r},
\end{equation}
which is a vector valued, zero-mean normally distributed white noise process with covariance (see \cite{Scerri2009} for proof)
\begin{equation}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into equation \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is given by
\begin{equation}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(\mathbf{r}_i - \mathbf{r}')\boldsymbol{\mu}_j(\mathbf{r}') \, d\mathbf{r}'.
\end{equation}
This completes derivation of the state-space representation of the multiresolution approximation of the neural field model. 
% \parham{Mike you can add state-space derivation using Galerkin projection here.}
